---
title: "Part 0: Installation Instructions"
pagetitle: "Part 0: Installation Instructions"
subtitle: "OCTRI-BERD Workshop July 2025"
author: "Jessica Minnier, Meike Niederhausen"
institute: "OHSU-PSU School of Public Health"
date: "7/17/25"
# format: 
#   revealjs:
#       incremental: false
#       scrollable: true
#       chalkboard: true
#       theme: [../sky_modified_smaller_font.scss]
#       width:  1050 #default 1050
#       height: 700 #default 700
#       slide-number: true
#       html-math-method: mathjax
html:
    link-external-newwindow: true
    toc: true
    embed-resources: true
execute:
  echo: true
  freeze: auto  # re-render only when source changes
# editor: visual
editor_options: 
  chunk_output_type: console
---

# Install recent versions of R and RStudio

Before we begin, please ensure you have the following installed:

1.  **R:** A recent version of R is required. You can download it from the [Comprehensive R Archive Network (CRAN)](https://cran.r-project.org/).
2.  **RStudio:** Download the latest version from the [RStudio website](https://posit.co/download/rstudio-desktop/). Quarto is built right into RStudio, making it the easiest way to get started. RStudio v2022.07 and later includes support for editing and preview of Quarto documents (the documentation below assumes you are using this build or a later version). If you are using Quarto within RStudio it is strongly recommended that you use the latest release of RStudio.

# Install R packages

You will also need to install a few R packages. Open RStudio and run the following code in the console:

```{r}
#| eval: false
install.packages(c("palmerpenguins", "tidyverse", "gt", "countdown"))
```

- **palmerpenguins**: For example datasets.
- **tidyverse**: A collection of R packages for data science.
- **gt**: For creating beautiful tables.
- **countdown**: To add timers to your slides.

If you do not already have **rmarkdown** installed, please install that as well:
```{r}
#| eval: false
install.packages("rmarkdown")
```

## Need an R & RStudio refresher?

* This workshop is intended for people with already some familiarity with R and RStudio. 
    * However, we will not be focusing on R coding. We will briefly talk about R code chunks and how to create and run them, but only with basic R code. 
* We are expecting basic familiarity with how to use RStudio, installing packages, and ideally also using projects. 
* If you would like refreshers on these topics, check out slides from our most recent __[Getting Started in R and Rstudio](https://berd-r-intro2024.netlify.app/slides_r_berd_2024#/title-slide)__ workshop from Sept 19, 2024:
    * [Intro to R and RStudio](https://berd-r-intro2024.netlify.app/slides_r_berd_2024#/introduction-to-r): slides 4-8
    * [R packages](https://berd-r-intro2024.netlify.app/slides_r_berd_2024#/r-packages-install-load): slides 23-29
    * [R projects](https://niederhausen.github.io/BSTA_511_F24/resources/Projects_in_R.html) - from BSTA 511 Fall 2024 class webpage


# Publishing webpages online

* If you would like to publish your website online, the quickest way is to use [Quarto Pub](https://quartopub.com/). Use the link to sign up for a free account. 
    * We will also provide instructions for GitHub website hosting, but not cover this more involved process during the workshop.
* For inspiration, you can also check out the [Quarto Gallery](https://quarto.org/docs/gallery/).


